---
title: "White Wine Quality by Jaekang Lee"
output:
  html_document:
    df_print: paged
---
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
#install.packages('GGally')
library(ggplot2)
library(dplyr)
library(tidyverse)
library(GGally)
library(ggplot2)
set.seed(42)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd('D:/Project/R Scripts/Exploratory-Data-Analysis/Udacity+Facebook')
wine <- read.csv('wineQualityWhites.csv')
```

The white wine quality dataset consists of 13 variables, with 4898 observations. Let's see what makes the best white wine!

```{r echo=FALSE, Univariate_Plots}
summary(wine)
ggplot(data=wine, 
       aes(x=quality))+ geom_bar(color='black' ,fill='pink') + scale_y_log10()
```

# Univariate Plots Section
*Note that we are most interested in 'quality' column as we want to see which variables help us find the best wine.

*It looks like Majority of our wines are quality between '5' to '7'. There are less than 10 of the highest quality '9' and dataset is normally distributed.

*Let's look at some boring bar plots of rest of the variables.

```{r echo=FALSE, warning=FALSE}
#Density
ggplot(data=subset(wine,density< quantile(density, 1)), 
       aes(x=density))+ geom_histogram(binwidth=0.0005, color='black' ,fill='pink')
#Alcohol
ggplot(data=wine, aes(x=alcohol))+ geom_histogram(binwidth=0.1, color='black' ,fill='pink')
#Sugar
ggplot(data=wine, aes(x=residual.sugar))+ 
  geom_histogram(binwidth=0.05, color='black' ,fill='pink') + scale_x_log10()
#Fixed * Volatile * Citric acid
wine$total_acid <- wine$fixed.acidity*wine$volatile.acidity*wine$citric.acid
ggplot(data=subset(wine, !is.na(total_acid)), aes(x=total_acid))+ 
  geom_histogram(binwidth=0.05,color='black' ,fill='pink') + scale_x_log10()
#Free sulfur dioxide rate
wine$free_rate <- wine$free.sulfur.dioxide/wine$total.sulfur.dioxide
ggplot(data=wine, aes(x=free_rate))+ geom_histogram(color='black' ,fill='pink')
#pH 
ggplot(data=wine, aes(x=pH))+ geom_histogram(binwidth=0.05,color='black' ,fill='pink')
#Residual
ggplot(data=wine, aes(x=chlorides))+ geom_histogram(binwidth=0.03,color='black' ,fill='pink') +
  scale_x_log10()
```
Boring as expected.

### Some interesting points:
* Since the dataset consists of real experts (human in the end), I think majority of their points comes from sweetness(sugar), acidity, alcohol and density.
* I've created total_acidity variable which doesn't make any sense but simplifies fixed, volatile and citric acids and free_rate which is proportion of free sulfur dioxide in total sulfur dioxide. Both for small dimension reduction purpose.
* Sugar had two main peaks at around 2 and 10. I guess majority of winemakers like to make their wine either sweet or not sweet. Most datas were skewed due to very few extreme outliers. 

# One shot look at all bivariates 

```{r echo=FALSE, Bivariate_Plots}
wine_subset <- wine[, c(5,6,9,10,11,12,13,14,15)]
names(wine_subset)
ggpairs(wine_subset[sample.int(nrow(wine_subset), 1000), ])
```

We can see that most scatterplots tends to form clusters with some outliers. Most interesting findings are...
*1.* Quality is most correlated with alcohol, density, chlorides
*2.* Alcohol is highly correlated with chloride(-0.82) and sugar(-0.46)
*3.* Sugar, alcohol and density are highly correlated 

Let's have a closer look.

```{r echo=FALSE, warning=FALSE}
wine$quality.bucket <- cut(wine$quality, breaks=c(2,4,7,9))
levels(wine$quality.bucket)
levels(wine$quality.bucket) <- c("Low Quality", "Medium Quality", "High Quality")
#BOX PLOTS
# ALCOHOL
ggplot(data=wine, 
       aes(y=alcohol)) +
  geom_boxplot() + 
  facet_wrap((~quality.bucket)) +
  scale_y_log10()
# DENSITY
ggplot(data=subset(wine,density<quantile(density,0.99)), 
       aes(y=density)) +
  geom_boxplot() + 
  facet_wrap((~quality.bucket)) +
  scale_y_log10()
#CHLORIDES
ggplot(data=subset(wine, chlorides < quantile(chlorides, 0.95) & 
                     chlorides > quantile(chlorides, 0.01)), 
       aes(y=chlorides)) +
  geom_boxplot() + 
  facet_wrap((~quality.bucket)) +
  scale_y_log10()
```

### Boxplots of different qualities
So clearly, high quality wines tend to have higher alcohol, lower density and lower chlorides than other qualities of wine. 

```{r echo=FALSE, warning=FALSE}
wine.wineByQuality <- wine %>%
  group_by(quality) %>%
  summarise(med_alcohol=median(as.numeric(alcohol)),
                           med_desity=median(as.numeric(density)),
                           med_chloride=median(as.numeric(chlorides)),
                           n=n()) %>%
  ungroup() %>%
  arrange(quality)
wine.wineByQuality
```

We used median here since most variables were severly skewed. We can somewhat guess typical characteristics of particular qualities in this chart. We can also see clear difference between quality 3 and 9 instantly. 

# Multivariate Plots Section

```{r echo=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(data=subset(wine,density < quantile(density,0.99)), 
      aes(x=(alcohol), y=density, color=residual.sugar)) +
   scale_y_log10(limits=c(0.987, 1.001)) +
  facet_wrap(~quality.bucket) +
  geom_point(alpha = 0.8, shape=3,size = 1) +
  scale_color_gradient(low="blue", high="pink")
```

### Multivariate Section
There's a lot going on here. First we can easily confirm that indeed density, alcohol and sugar closely related.

*1.* Higher sugar levels cover upper right region while lower sugar level covers lower left region. This means higher density and alcohol, higher the sugar level and vice versa. 

*2.* Alcohol and density are highly correlated. I suspect this is because alcohol is main source of reducing density since every other ingredients are likely to increase density. 

*3.* High quality wines have a cluster at the higher alcohol and lower density quadrant. In contrast, low quality wines have a cluster located further towards 1st and 3rd quadrants. Medium quality wines covers both area. 

*4.* No wines seem to exist lowest alcohol and lowest density region.

# Multivariate Linear Regression

```{r echo=FALSE, warning=FALSE, RandomForestClassifier}
model <- lm(quality ~ density + alcohol + chlorides, data = wine)
summary(model)
```

Of course this is just a old textbook solution. It is terrible and very time consuming to try different variable combinations. Lets use random forest instead.

# Random Forest Classifier on all quality [3,9]

```{r echo=FALSE, warning=FALSE}
#install.packages('randomForest')
library(randomForest)
#install.packages('caret')
library(caret)
na.omit(wine_subset)
wine_subset_subset <- wine[,c(2,3,4,5,6,7,8,9,10,11,12,13,14,15)]
preproc2 <- preProcess(wine_subset_subset[,c(1:11,13:14)],method=c("range"))
norm2 <- predict(preproc2, wine_subset_subset[,c(1:11,13:14)])
norm2$quality <- wine_subset_subset$quality
## 80% of the sample size
smp_size <- floor(0.8 * nrow(norm2))

## set the seed to make your partition reproducible
set.seed(42)

train_ind <- sample(seq_len(nrow(norm2)), smp_size)
train <- norm2[train_ind, ]
test <- norm2[-train_ind, ]
rf_classifier1= randomForest(factor(quality)~., data=train, ntree=500, importance=TRUE)
rf_classifier1
```

We got an accuracy of 69.4%. It is probably hard even for experts to 
distinguish quality level difference of 1 perfectly. 

# RFC(All quality) Interpretation (MeanDecreaseAccuracy, MeanDecreaseGini)

```{r echo=FALSE, warning=FALSE}
varImpPlot(rf_classifier1)
```

* To explain a little about the plots, MeanDecreaseAccuracy tells us how bad our predictions become if we were to omit that variable completly. MeanDecreaseGini tells us how clean the splits are if that variable was used to split the data.
* This had me wonder if I bucket some quality together into only 3 qualities, 'High', 'Medium', and 'Low' I should be able to increase accuracy by a significant amount.

# Random Forest Classifier on bucketed quality (High, Med, Low)

```{r echo=FALSE, warning=FALSE}
wine_subset_subset <- wine[,c(2,3,4,5,6,7,8,9,10,11,12,13,14,15)]
preproc2 <- preProcess(wine_subset_subset[,c(1:11,13:14)],method=c("range"))
norm2 <- predict(preproc2, wine_subset_subset[,c(1:11,13:14)])
norm2$quality <- wine_subset_subset$quality
norm2$quality.bucket <- cut(norm2$quality, breaks=c(2,4,7,9))
levels(norm2$quality.bucket) <- c("Low Qual.", "Med. Qual.", "High Qual.")
norm2 <- norm2[-c(14)]
## 80% of the sample size
smp_size <- floor(0.8 * nrow(norm2))

## set the seed to make your partition reproducible
set.seed(42)

train_ind <- sample(seq_len(nrow(norm2)), smp_size)
train <- norm2[train_ind, ]
test <- norm2[-train_ind, ]
rf_classifier= randomForest(factor(quality.bucket)~., data=train, ntree=500, importance=TRUE)
rf_classifier
```

Ok we've achieved 94.46% accuracy.

# RFC(bucketed) Interpretation (MeanDecreaseAccuracy, MeanDecreaseGini)
```{r echo=FALSE, warning=FALSE}
varImpPlot(rf_classifier)
```
```{r echo=FALSE, warning=FALSE}
predictions <- predict(rf_classifier,test)
table(test[,c(14)],predicted=predictions)
```
93.82% accuracy on our test set! So it looks like there barely any overfitting and it generalizes really well too. We can also note that alcohol once again is most important by a significant amount. Followed by many acids and density & sugar giving little meanings.

# RFC(bucketed) Interpretation 2 (ROC Curve)
###https://www.blopig.com/blog/2017/04/a-very-basic-introduction-to-random-forests-using-r/  (Reference)
```{r echo=FALSE, warning=FALSE}
#install.packages('ROCR')
library(ROCR)

predictions <- predict(rf_classifier,test,type="prob")
colours <- c("#F8766D","#00BA38","#619CFF")
classes <- levels(test$quality.bucket)
# For each class
for (i in 1:3)
{
 true_values <- ifelse(test[,c(14)]==classes[i],1,0)
 pred <- prediction(predictions[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve",col=colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve",col=colours[i],add=TRUE) 
 }
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}
```

Red - Low Qual., Green - Med Qual., Blue - High Qual.
Here we can see that model is very good at picking out the low quality wines but have harder time picking out high quality wines.In fact, if you recall the 7 level quality model, that model didn't predict any sample to be high quality. So we can conclude that bad wines are easy to distinguish but best wines are hard to detect.

# Reflection
The white wine set contains information on almost 5000 wine data which all have been evaluated by at least 3 different wine experts. We explored many variables and predicted that since density, chlorides, and alcohol had high correlation, their combination will define the quality of any wine. We also looked at an interesting relationship between chemicals like sugar makes density go up and alcohol go down. 

Our random forest quantifier agreed with alcohol being important but disagreed with density and chloride being important. Instead it told us a lot of acids which had little to no correlation were very important. And proceeded to succesfully predict 94.46% on training set and 93.82% on test set. Also note it had harder time distinguishing higher quality wines than lower quality wines by looking at its MeanDecreaseAccuracy. Meaning other factors such as presentation or color may may be necessary divide the 'best wine' from good wines.

In conclusion, as predicted from my idea; acidity, sweetness, alcohol level but density plays a very important factor in determining high quality white wines. So although I've never had wine in my life, I can taste one and tell if it is good or not : )
